---
title: "Classification of Acute Respiratory Distress Syndrome"
author: "Robert Edwards"
output:
  pdf_document:
    latex_engine: pdflatex
    number_sections: no
    fig_caption: yes
    
  word_document: default
  
  html_document:
    code_folding: hide
    df_print: paged
    fig_caption: yes
    
geometry: "left=3cm,right=3cm,top=3cm,bottom=3cm"
header-includes: 
  \usepackage[bottom]{footmisc}
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{color}
  \usepackage{xcolor}
---

```{r setup, include=FALSE, echo = FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
options()
```

```{r libraries, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(moderndive)
library(skimr)
library(tidyr)
library(kableExtra)
library(gridExtra)
library(xtable)
library(knitr)
library(GGally)
library(broom)
library(naniar) # missing data eda
library(VIM) # data imputation
```

```{r import_data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
data.raw.df <- read.csv(file = "../data/ARDSdata.csv", header = TRUE)
```


```{r reduce, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
Day1_vars <- c("Hospital_Survival", 
               "Duration_ECMO",
               "Day1ECMO_RR", 
               "Day1ECMO_Vt", 
               "Day1ECMO_FiO2", 
               "Day1ECMO_Ppeak", 
               "Day1ECMO_Pmean", 
               "Day1ECMO_PEEP", 
               "Day1ECMO_PF", 
               "Day1ECMO_SpO2", 
               "Day1ECMO_PaCO2", 
               "Day1ECMO_pH", 
               "Day1ECMO_BE", 
               "Day1ECMO_Lactate", 
               "Day1ECMO_NAdose", 
               "Day1ECMO_MAP", 
               "Day1ECMO_Creatinine", 
               "Day1ECMO_Urea", 
               "Day1ECMO_CK", 
               "Day1ECMO_Bilirubin", 
               "Day1ECMO_Albumin", 
               "Day1ECMO_CRP", 
               "Day1ECMO_Fibrinogen", 
               "Day1ECMO_Ddimer", 
               "Day1ECMO_ATIII", 
               "Day1ECMO_Leukocytes", 
               "Day1ECMO_Platelets", 
               "Day1ECMO_TNFa", 
               "Day1ECMO_IL6", 
               "Day1ECMO_IL8", 
               "Day1ECMO_siIL2")
data.df <- data.raw.df[, -which(names(data.raw.df) %in% Day1_vars)]
```


```{r factor, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE, results="hide"}
# Select the first 450 rows.  The others are blank
data.df <- data.df %>%
  slice(1:450) 
#  suppressMessages()  # suppress output
```

```{r factor, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE, results="hide"}
# Make categorical variables factors
data.df <- data.df %>%
  mutate(Indication = factor(Indication)) %>% 
  mutate(ECMO_Survival = factor(ECMO_Survival)) %>%
#  mutate(Hospital_Survival = factor(Hospital_Survival)) %>%
  mutate(Gender = factor(Gender)) %>%
  suppressMessages()  # suppress output
```


```{r missing_data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
vis_miss(data.df)
```

An upset plot from the `UpSetR` package can be used to visualise the patterns of missingness, or rather the combinations of missingness across cases. To see combinations of missingness and intersections of missingness amongst variables, use the `gg_miss_upset` function:

```{r missing_data_patterns, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
gg_miss_upset(data.df)
```

If there are 40 intersections, there will be up to 40 combinations of variables explored. The number of sets and intersections can be changed by passing arguments `nsets = 10` to look at 10 sets of variables, and nintersects = 50 to look at 50 intersections.

```{r num_missing, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
#n_var_miss(data.df)
gg_miss_upset(data.df, nsets = n_var_miss(data.df)/3)
```

This plot shows the number of missing values in each variable in a dataset. It is powered by the `miss_var_summary()` function.  
```{r var_missing, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
gg_miss_var(data.df, show_pct = TRUE) +
  labs("% Data Missing in Biomarker Data")
```

This plot shows the number of missings in each column, broken down by a categorical variable from the dataset. It is powered by a `dplyr::group_by` statement followed by `miss_var_summary()`.

```{r missing_fct, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
gg_miss_fct(x = data.df, fct = Indication) + 
  labs("NAs in Biomarkers by Age")
```

This plot shows the cumulative sum of missing values, reading the rows of the dataset from the top to bottom. It is powered by the `miss_case_cumsum()` function.

```{r cum_missing, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
gg_miss_case_cumsum(data.df)
```


This plot shows the cumulative sum of missing values, reading columns from the left to the right of your dataframe. It is powered by the `miss_var_cumsum()` function.

```{r cum_var_missing, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
gg_miss_var_cumsum(data.df)
```



#Exploratory Data Analysis
```{r boxplot, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
n <- ncol(data.df)
feature_names <- colnames(data.df[,7:n])
data_standard.df <- data.df %>%
  select(7:n) %>%
  mutate_each_(funs(scale), vars=feature_names) %>%
  glimpse()

data_standard.df %>% 
#  select(-loudness, -popularity) %>% 
  gather(key = Feature, value = Value) %>% 
  ggplot(aes(x = Feature, y = Value, fill = Feature)) +
#  ggplot(stack(track_features.df), aes(x = ind, y = values)) +
    geom_boxplot() +
    xlab("Features") + 
    ylab("Average") + 
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position="none",
          axis.text.x = element_text(angle = 60, hjust = 1)) + 
    scale_fill_brewer(palette="Spectral")
```


#Imputation
Here I will try various methods for data imputation.

##I. KNN Imputation

```{r knn_imputation, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
data_impute.df <- kNN(data.df)

data_impute.df %>%
  glimpse()
```

```{r standardizing, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
# Scaling the data and logging it
feature_names <- colnames(data_impute.df[,7:36])
data_standard.df <- data_impute.df %>%
  select(7:36) %>%
  mutate_each_(funs(scale), vars=feature_names) %>%
  mutate_each_(funs(log), vars=feature_names) %>%
  glimpse()

data_standard.df %>% 
  gather(key = Feature, value = Value) %>% 
  ggplot(aes(x = Feature, y = Value, fill = Feature)) +
#  ggplot(stack(track_features.df), aes(x = ind, y = values)) +
    geom_boxplot() +
    xlab("Features") + 
    ylab("Average") + 
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position="none",
          axis.text.x = element_text(angle = 60, hjust = 1)) + 
    scale_fill_brewer(palette="Spectral")
```


```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
data_standard.df %>% 
  gather(key = Feature, value = Value) %>% 
  ggplot(aes(x=Value, color=Feature)) +
      geom_density(alpha=0.5) +
      geom_vline(aes(xintercept=0)) +
      theme_bw() +
      scale_fill_brewer(palette="Spectral") 
```




